---
title: "p8105_hw6_af3341"
author: "Alana Ferris"
date: "2022-11-22"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(modelr)
library(mgcv)
library(patchwork)

set.seed(1)

knitr::opts_chunk$set(
  fig.width = 7,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


# Problem 1

## Reading in the data
```{r, message = FALSE, warning = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## 5000 bootstrap samples
```{r, message = FALSE, warning = FALSE}
weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  ggplot(aes(x = r.squared)) + geom_density()
```

The 5000 bootstrap sample distribution appears left skewed. This implies that the majority of 5000 bootstrap sample $R^2$ values are closer to 1, which indicates our model is a good fit for this data set. 

## Reading in the data
```{r, message = FALSE, warning = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## 5000 bootstrap samples
```{r, message = FALSE, warning = FALSE}
weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  ggplot(aes(x = r.squared)) + geom_density()
```

The 5000 bootstrap sample distribution appears left skewed. This implies that the majority of 5000 bootstrap sample $R^2$ values are closer to 1, which indicates our model is a good fit for this data set. 

## 95% confidence interval for the 5000 bootstrap samples 
```{r, message = FALSE, warning = FALSE}
weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::tidy)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  select(id = `.id`, term, estimate) %>% 
  pivot_wider(
    names_from = term, 
    values_from = estimate) %>% 
  rename(beta0 = `(Intercept)`, beta1 = tmin) %>% 
  mutate(log_b0b1 = log(beta0 * beta1)) %>% 
  ggplot(aes(x = log_b0b1)) + geom_density()
```

# Problem 2

## Reading in and cleaning homicide data

```{r, message = FALSE, warning = FALSE}
homicides = 
  read_csv("./data/homicide-data.csv")

total_homicide = 
  homicides %>% 
  mutate(city = str_replace(city, "$", ", ")) %>% 
  mutate(city_state = 
    paste0(city, state)) %>% 
  mutate(unsolved = 
           case_when(disposition == 'Closed without arrest' | 
                    disposition == 'Open/No arrest' ~ TRUE, 
                    disposition == 'Closed by arrest'~ FALSE)) %>% 
  mutate(
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White"), 
    victim_sex = fct_relevel(victim_sex, "Male")) %>% 
  filter(city_state != "Dallas, TX" && 
         city_state != "Pheonix, AZ" && 
         city_state != "Kansas City, MO" && 
         city_state != "Tulsa, AL") %>% ## for some reason it isn't dropping tulsa, AL
  filter(victim_race == "White" | victim_race == "Black") %>% 
  group_by(city_state)

```

## Fitting glm function to Baltimore

```{r, message = FALSE, warning = FALSE}
balt_df = 
  total_homicide %>% 
  filter(city_state == "Baltimore, MD") %>% 
  select(unsolved, victim_age, victim_race, victim_sex)

fit_logistic = 
  balt_df %>%
  glm(unsolved ~ victim_age + victim_race + victim_sex, 
      data = ., 
      family = binomial()) # is this keeping the other variables fixed, problem only interested in est and CI of OR of male and female 
  
 fit_logistic %>%  
   broom::tidy() %>% 
   mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(term, OR, starts_with("CI")) %>% 
  knitr::kable(digits = 3)
```

Homicides in which the victim is Black increases odds of the homicide being unresolved by about 2x compared to victims who are White. Female victims have about a 60% reduced risk of their case being unsolved. There is no association between a victim's age and the odds that their homicide is unsolved.   

## Mapping glm across all cities 

```{r, message = FALSE, warning = FALSE}
all_cities =
  total_homicide %>% 
  select(unsolved, victim_age, victim_race, victim_sex) %>% 
  filter(city_state != "Tulsa, AL")

all_nested = 
  all_cities %>% 
  nest(data = unsolved:victim_sex)

all_unnested = 
  all_nested %>% 
  mutate(
    glm_cities = map(.x = data, ~glm(unsolved ~ victim_age + victim_race + victim_sex, 
                                     data = .x, 
                                     family = binomial())),
    results = map(glm_cities, broom::tidy)
  ) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, OR, starts_with("CI")) %>% 
  filter(term == "victim_sexFemale")

```

## Plot of estimated ORs and CIs for each city

```{r, message = FALSE, warning = FALSE}
all_unnested %>% 
  ggplot(aes(x = reorder(city_state, -OR), y = OR)) +
  geom_point() + 
  geom_errorbar(aes(x = city_state, ymin = OR - CI_lower, ymax = OR + CI_upper)) + 
  theme(axis.text.x = element_text(angle = 65, hjust = 1, size = 6)) +
   labs(
    title = "Odds of a Homicide Being Unsolved Based on Victim Sex Across Cities",
    x = "U.S. City",
    y = "Odds Ratio")
```

On average, there is a higher odds of a homicide being unsolved if the victim is male which we can see because the majority of the odds ratios for female victims' homicides being unsolved is below 1. 

# Problem 3 

## Reading in and cleaning birthweight data 

```{r, message = FALSE, warning = FALSE}
birthweight = read_csv('./data/birthweight.csv') %>% ## remind rae to do ./ 
  mutate(
    babysex = recode(babysex, `1` = "male", `2` = "female"), 
    babysex = as.factor(babysex), 
    frace = recode(frace, `1` = "white", `2` = "black", `3` = "asian", `4` = "puerto rican", `8` = "other", `9` = "unknown"), 
    frace = as.factor(frace), 
    malform = recode(malform, `0` = "absent", `1` = "present"), 
    malform = as.factor(malform), 
    mrace = recode(mrace, `1` = "white", `2` = "black", `3` = "asian", `4` = "puerto rican", `8` = "other"), 
    mrace = as.factor(mrace)
  ) 
```

## Regression model for birthweight 

```{r, message = FALSE, warning = FALSE}
fit = birthweight %>% 
  mutate(
    mrace = fct_infreq(mrace),
    babysex = fct_infreq(babysex)) %>% 
    lm(bwt ~ mrace + smoken + babysex + gaweeks + parity + blength, data = .) 
  
fit %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 4)

predict_table = add_predictions(birthweight, fit) %>% 
  select(bwt, pred, everything())

modelr::add_residuals(predict_table, fit) %>% 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point()
```

When building the regression model, I included factors that have been shown to impact birthweight: mother's race, mother's smoking habits, number of live births the mother had prior to this pregnancy, baby's sex, gestational age, length.  

## 2 models for comparison

```{r, message = FALSE, warning = FALSE}
length_gsa = birthweight %>% 
  lm(bwt ~ blength + gaweeks, data = .)

multi = birthweight %>% 
  lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = .)

birthweight %>% 
  gather_predictions(fit, length_gsa, multi) %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = blength, y = bwt)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red") + 
  facet_wrap(~model)
```

## Cross validation 

```{r, message = FALSE, warning = FALSE}
cv_df = 
  crossv_mc(birthweight, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

cv_mutated = 
  cv_df %>% 
  mutate(
    fit = map(train, ~lm(bwt ~ mrace + smoken + babysex + gaweeks + parity + blength, data = .x)),
    length_gsa = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    multi = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = .x))) %>% 
  mutate(
    rmse_fit = map2_dbl(fit, test, ~rmse(model = .x, data = .y)),
    rmse_length_gsa = map2_dbl(length_gsa, test, ~rmse(model = .x, data = .y)),
    rmse_multi = map2_dbl(multi, test, ~rmse(model = .x, data = .y)))

cv_mutated %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

Our model seems reasonably fit in comparison to the other two models. However, the multi-fit model has the lowest RMSE, so is technically the best fit. 